Maximizing Ink in Symmetric Partial Edge Drawings of $k$-plane Graphs

### 2018-01-09 eurocg/paper.tex
verwaisten satz entfernt
merge

Section~\ref{sec:hardness} showed that \maxsped is generally \NP-hard for $k \ge 3$. In this section we consider the special case that the intersection graph of the $k$-plane input drawing is a tree or a cactus. In both cases we present polynomial-time dynamic programming algorithms. Let $ C = (V,E) $ be the intersection graph of a given drawing $\Gamma$ of graph $ G $ as defined in Section~\ref{sec:preliminaries}. Let $ u \in V $ and $ \deg(u) = \delta $. 
Then for the corresponding segment $ s(u) \in S $ there are $ \delta + 1 $ possible stub pairs including drawing the whole segment, see Fig.~\ref{fig:tree}. We store the lengths of the possible stub pairs as node weights $ \ell_1(u), \dots, \ell_\delta(u) \in \mathbb{R}_+$ and assume that $ \ell_1(u),\dots,\ell_\delta(u) $ are sorted from shorter to longer stubs. We define  $ \ell_0(u) $ as the length of the whole segment $ s(u) $.

\begin{figure}[tbp]
	\centering
 	\includegraphics{tree_illustration}
	\caption{A segment with  intersecting segments and the different induced stub lengths.} % The shaded areas are the stubs, $ p(u) $ is the parent of $ u $. The stubs in the boxed area contribute to $ \solshort(u) $.}
	\label{fig:tree}
\end{figure}

@@

\subsection{Trees}
\label{sec:tree}

@@

Figure~\ref{fig:tree} highlights the stub lengths that contribute to $\solshort(u)$. We recursively define % for a vertex $ u $ and stub length $ \ell_i(u) $:

@@

		\sollong(v) & \text{otherwise}
	\end{cases}.

@@

So by solving Recurrence~(\ref{rec:tree}) in $O(mk)$ time we find an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ using standard backtracking from $\max\{T_0(r), \dots, T_{\deg(r)}(r)\}$ for the root $r$ of $C$.	
Since the intersection graph $C$ is a tree with $O(m)$ edges it can be computed in $O(m \log m)$ time and we obtain the following theorem.

\begin{theorem}
	\label{thm:tree}
	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $. If the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $ problem \maxsped can be solved in $ O(mk + m \log m)) $ time and space.
\end{theorem}
% \begin{proof}
% 	Since $ C $ is known to be a tree, and hence $ |E| = O(m) $, we can compute the intersection graph and the stub lengths in $ O(m\log m) $ time and space.
% 	With Lemma~\ref{lem:compute_vertex} and Recurrence~\ref{rec:tree} we need one pass through the tree to find the roots maximum in $ O(mk) $ time. Adding the bounds yields the desired runtime of $ O(mk + m \log m)) $.
% %	To solve the \maxsped problem on $ G $ we need to compute $ \max \{T_i(r) \mid 0 \leq i \leq k \} $ for the root $ r \in V $ of $ C $ as seen above. Doing this bottom up requires one pass through $ C $ hence by applying Lemma~\ref{lem:compute_vertex} we need $ O(mk) $ time to compute all $ T_i(u) $ for all $ u \in V $. For every vertex we only need to compute $ O(k) $ values and store two of them, namely $ \sollong(u) $ and $ \solshort(u) $, so we do not need more than $ O(mk) $ memory. Adding the bounds yields the desired runtime of $ O(mk + m \log m)) $.
% %
%
%
%
% 	Retrieving the actual stub length for each edge in $ G $ needed to produce the SPED can be done by backtracking the subsolutions which requires at most $ O(mk) $ time.
% \end{proof}

@@

Let $ C = (V,E) $ be a cactus graph and moreover the intersection graph of a graph $ G $ and straight-line drawing $ \Gamma $. Additionally let $ \mathcal T = (\mathcal V, \mathcal E) $ be the block-cut tree of $ C $ we define for each vertex $ u \in \mathcal V $ two values $ a,b \in \mathbb{R}_+ $, w.l.o.g $ a \leq b $. If $ u $ is an articulation vertex this are exactly the positions where the cycle represented by $ p(u) $ intersects $ s(A(u)) $ in $ \Gamma $. If $ u $ is a block vertex $ a,b $ are the points where edges $ s(v), v \in \textit{cycle}(u) $ intersect $ s(A(p(u))) $. Similar to Section~\ref{sec:tree} we define three values for every vertex $ u \in \mathcal V $ with degree $ \delta $:
\begin{align*}
	\solshort(u) &= \max\{T_0(u),\dots,T_\delta(u)\}\\
	\solmid(u) &= \max\{T_1(u),\dots,T_{a}(u)\}\\
	\sollong(u) &= \max\{T_1(u),\dots,T_{b}(u)\}
\end{align*}
For an articulation vertex $ u \in \mathcal V $:
\begin{align}
\label{rec:cactus_articulation}
T_i(u) = \ell_i(A(u)) + \sum_{v \in c(u)}
\begin{cases}
\solshort(v) & \text{if } \ell_i(A(u)) \leq a \\
\solmid(v) & \text{if } a < \ell_i(A(u)) \leq b \\
\sollong(v) & \text{else}
\end{cases}
\end{align}
	Let $ G $, $ \Gamma $,  $ C = (V,E) $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above and $ u \in \mathcal V $ a leaf of $ \mathcal T$. Then either $ u $ is an articulation vertex and to be treated as in Lemma~\ref{lem:compute_vertex} or a cycle. In the latter case we find the optimal solution to the \maxsped problem on $ C[\textit{cycle}(u)] $ with the algorithm from Bruckdorfer et al. \cite{bcgkmn-pped-17}. Using the above described method we derive $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $, this are exactly the optimal values depending on how the cycle interacts with the articulation vertex $ p(u) $.

>>>>

Section~\ref{sec:hardness} showed that \maxsped is generally \NP-hard for $k \ge 3$. In this section we consider the special case that the intersection graph of the $k$-plane input drawing is a tree or a cactus. In both cases we present polynomial-time dynamic programming algorithms. Let $ C = (V,E) $ be the intersection graph of a given drawing $\Gamma$ of graph $ G $ as defined in Section~\ref{sec:preliminaries}. Let $ u \in V $ and $ \deg(u) = \delta $. 
Then for the corresponding segment $ s(u) \in S $ there are $ \delta + 1 $ possible stub pairs including drawing the whole segment. We store the lengths of the possible stub pairs as node weights $ \ell_1(u), \dots, \ell_\delta(u) \in \mathbb{R}_+$ and assume that $ \ell_1(u),\dots,\ell_\delta(u) $ are sorted from shorter to longer stubs. We define  $ \ell_0(u) $ as the length of the whole segment $ s(u) $.

@@

\subsection{Trees}
\label{sec:tree}
\begin{figure}[tbp]
	\centering
 	\includegraphics{tree_illustration}
	\caption{One edge $ s(u) $ with intersecting segments. The shaded areas are the stubs, $ p(u) $ is the parent of $ u $. The stubs in the boxed area contribute to $ \solshort(u) $.}
	\label{fig:tree}
\end{figure}

@@

Based on these two values we can recursively define $ T_i(u) $: % for a vertex $ u $ and stub length $ \ell_i(u) $:

@@

		\sollong(v) & \text{else}
	\end{cases}

@@

\begin{theorem}
	\label{thm:tree}
	Let $ G $ be a simple graph with $m$ edges and $ \Gamma $ a straight-line drawing of $ G $. If the intersection graph $ C = (V,E) $ of $ G $ is a tree with maximum degree $ k \in \mathbb{N} $ the \maxsped problem can be solved in $ O(mk + m \log m)) $ time and space.
\end{theorem}
\begin{proof}
	Since $ C $ is known to be a tree, and hence $ |E| = O(m) $, we can compute the intersection graph and the stub lengths in $ O(m\log m) $ time and space.
	With Lemma~\ref{lem:compute_vertex} and Recurrence~\ref{rec:tree} we need one pass through the tree to find the roots maximum in $ O(mk) $ time. Adding the bounds yields the desired runtime of $ O(mk + m \log m)) $.
%	To solve the \maxsped problem on $ G $ we need to compute $ \max \{T_i(r) \mid 0 \leq i \leq k \} $ for the root $ r \in V $ of $ C $ as seen above. Doing this bottom up requires one pass through $ C $ hence by applying Lemma~\ref{lem:compute_vertex} we need $ O(mk) $ time to compute all $ T_i(u) $ for all $ u \in V $. For every vertex we only need to compute $ O(k) $ values and store two of them, namely $ \sollong(u) $ and $ \solshort(u) $, so we do not need more than $ O(mk) $ memory. Adding the bounds yields the desired runtime of $ O(mk + m \log m)) $.
%

By solving Recurrence~(\ref{rec:tree}) we find an optimal solution to the \maxsped problem on $ G $ with drawing $ \Gamma $ as $\max\{T_0(r), \dots, T_{\deg(r)}(r)\}$ for the root $r$ of $C$.

	Retrieving the actual stub length for each edge in $ G $ needed to produce the SPED can be done by backtracking the subsolutions which requires at most $ O(mk) $ time.
\end{proof}

@@

Let $ C = (V,E) $ be a cactus graph and moreover the intersection graph of a graph $ G $ and straight-line drawing $ \Gamma $. Additionally let $ \mathcal T = (\mathcal V, \mathcal E) $ be the block-cut tree of $ C $ we define for each vertex $ u \in \mathcal V $ two values $ a,b \in \mathbb{R}_+ $, w.l.o.g $ a \leq b $. If $ u $ is an articulation vertex this are exactly the positions where the cycle represented by $ p(u) $ intersects $ s(A(u)) $ in $ \Gamma $. If $ u $ is a block vertex $ a,b $ are the points where edges $ s(v), v \in \textit{cycle}(u) $ intersect $ s(A(p(u))) $. Similar to Section~\ref{sec:tree} we define three values for every vertex $ u \in \mathcal V $ with degree $ \delta $:
\begin{align*}
	\solshort(u) = \max\{T_0(u),\dots,T_\delta(u)\}\\
	\solmid(u) = \max\{T_1(u),\dots,T_{a}(u)\}\\
	\sollong(u) = \max\{T_1(u),\dots,T_{b}(u)\}
\end{align*}
For an articulation vertex $ u \in \mathcal V $:
\begin{align}
\label{rec:cactus_articulation}
T_i(u) = \ell_i(A(u)) + \sum_{v \in c(u)}
\begin{cases}
\solshort(v) & \text{if } \ell_i(A(u)) \leq a \\
\solmid(v) & \text{if } a < \ell_i(A(u)) \leq b \\
\sollong(v) & \text{else}
\end{cases}
\end{align}
	Let $ G $, $ \Gamma $,  $ C = (V,E) $ and $ \mathcal T = (\mathcal V,\mathcal E) $ as above and $ u \in \mathcal V $ a leaf of $ \mathcal T$. Then either $ u $ is an articulation vertex and to be treated as in Lemma~\ref{lem:compute_vertex} or a cycle. In the latter case we find the optimal solution to the \maxsped problem on $ C[\textit{cycle}(u)] $ with the algorithm from Bruckdorfer et al. \cite{bcgkmn-pped-17}. Using the above described method we derive $ \solshort(u) $, $ \solmid(u) $ and $ \sollong(u) $, this are exactly the optimal values depending on how the cycle interacts with the articulation vertex $ p(u) $.
